Stochastic Wind Power Forecasting

### 2020-04-09 LaTeX_Code/Paper/paper_1line.tex
update Lamperti, likelihood

\begin{figure}[h]
  %\includegraphics[width=26mm,scale=1]{}
  \caption{Insert Figure showing the difference with and without derivative tracking}
  \label{fig:derivative_tracking}
\end{figure}

@@

Our model (\ref{VtSDE}) for the forecast error has a diffusion term that depends on the state variable $V_t$. To estimate the unknown model parameters, a recommended technique is to modify the SDE (\ref{VtSDE}) by applying the so-called Lamperti transform (see \cite[40--41]{iacus1}, \cite{moma}, \cite[98--100]{saso}) to the process $V$ to obtain a SDE for the transformed process whose diffusion term does not depend anymore on the state variable. To this purpose, we consider the following Lamperti transformation  
%{\color{red} (Remind that continuous-time models with different diffusion term are singular and the likelihood cannot be written down. Besides, it can be unstable to work with low frequency. If the frequency of the data is too low, the Lamperti transform helps to detect this problem (in principle, it is not clear how to determine the right frequency for estimation. )}
%\begin{equation}
%Z_t = h(V_t, t) = \frac{1}{\sqrt{2 \theta_0 \alpha}} \int_{\xi}^{x} \frac{1}{\sqrt{(u + p_t)(1 - u - p_t)}} \,du \Bigg{\vert}_{x = V_t} \,,
%\end{equation}
%where $\xi$ is an arbitrary point of the state space of the process $V$. The choice of $\xi = \frac{1}{2} - p_t$, where $p_t$ is a known input, 
\begin{align}
Z_t = h(V_t, t)  & = \frac{1}{\sqrt{2 \alpha \theta_0 }} \int \frac{1}{\sqrt{(v + p_t)(1 - v - p_t)}} \,dv \Bigg{\vert}_{v = V_t}  \nonumber \\
& = \frac{1}{\sqrt{2 \alpha \theta_0 }} \arcsin (2(V_t + p_t) -1)
\end{align}
that, after applying It\^{o}'s formula on $h(V_t, t)$, leads to the following SDE with state independent diffusion term
\begin{align}
dZ_t = \Bigg[  & \frac{\dot{p}_t}{ \sqrt{2 \alpha \theta_0 (V_t + p_t)(1 - V_t - p_t)}}  \nonumber \\ 
& + \frac{- \theta_t V_t}{ \sqrt{2  \alpha \theta_0 (V_t + p_t)(1 - V_t - p_t)}} - \frac{1}{4} \frac{\sqrt{2 \alpha \theta_0} \left( 1 - 2 (V_t + p_t)\right)}{\sqrt{(V_t + p_t)(1 - V_t - p_t)}}  \Bigg] \,dt + dW_t \,. \label{eq:stindepSDE}
\end{align}
After replacing $V_t = \frac{\sin(\sqrt{2 \alpha \theta_0} Z_t) + 1}{2} - p_t$ in (\ref{eq:stindepSDE}), we obtain that the  process $Z$ satisfies the SDE
\begin{align}
dZ_t & = \Bigg[ \frac{2 \dot{p}_t}{\sqrt{2 \alpha \theta_0} \cos{(\sqrt{2 \alpha \theta_0} Z_t)}} + \frac{- \theta_t (\sin(\sqrt{2 \alpha \theta_0} Z_t) + 1 - 2 p_t)}{\sqrt{2 \alpha \theta_0} \cos{(\sqrt{2 \alpha \theta_0} Z_t)}} \nonumber \\
&  + \frac{\sqrt{2 \alpha \theta_0} \sin{(\sqrt{2 \alpha \theta_0} Z_t)}}{2 \cos{(\sqrt{2 \alpha \theta_0} Z_t)}} \Bigg] \,dt + dW_t \nonumber \\
&  = \left[  \frac{(\alpha \theta_0 - \theta_t) \sin(\sqrt{2 \alpha \theta_0 } Z_t) - \theta_t (1 - 2 p_t) + 2  \dot{p}_t }{\sqrt{2 \alpha \theta_0} \cos{(\sqrt{2 \alpha \theta_0} Z_t)}}  \right] \,dt + dW_t \,. 
\end{align}

@@

Suppose that any of $M$ non-overlapping paths of the continuous-time It\^{o} process $V = \{ V_t, t  \in [0,T] \}$ is sampled at $N$ equispaced discrete points with given length interval $\Delta$, and let $ V^{M,N}=\{ V_{t_1^{M,N}} , V_{t_2^{M,N}} ,\ldots , V_{t_N^{M,N}} \}$ denote this random sample.

Let $\rho_i(v \vert v_{j, t_{i-1}} ; \bm{\theta})$ denote the conditional probability density of $V_{j, t_i}$ given $V_{j, t_{i-1}} = v_{j, t_{i-1}}$ evaluated at $v$, where $\bm{\theta} = (\theta_0, \alpha)$ are the unknown model parameters.

The It\^{o} process $V$ defined by the SDE (\ref{VtSDE}) is Markovian, then the likelihood function of the data can be written as the following product of transition densities:  
%Consider the set of $M$ paths with $N$ observations each, $ V^{M,N}=\{ V_{t_1^{M,N}} , V_{t_2^{M,N}} ,\ldots , V_{t_N^{M,N}} \}$, observed in $\Delta_N$ intervals.
\begin{equation}
\mathcal{L}(\bm{\theta},\delta; V^{M,N}) = \prod\limits_{j=1}^M \prod\limits_{i=1}^N \rho ( {V_{j, t_i}| V_{j, t_{i-1}}} ; p_{[t_{i-1}, t_{i}]},  \bm{\theta} )  \, \rho_0 (V_{j, t_0}|V_{j, t_{-\delta}};\bm{\theta},\delta) \,.
\label{likelihood}
\end{equation}

 {\color{red} In \ref{likelihood} modify the expression for $\rho_0$ contemplating the lag, say $\delta$, between the time when the forecast is done and the first time of forecasting. $\rho_0$ can be written as a conditional density between these two times. Once $\theta_0$ and $\alpha$ have been estimated, one could estimate $\delta$.}

@@

\begin{align}
\frac{ \partial f }{\partial t } & \rho_i(v_{j,i} ,t \vert v_{j,i-1} ,  t_{j,i-1} ; \bm{\theta} )= - \frac{\partial}{ \partial v} (- \theta_t v \, \rho_i(v ,t_{j,i} \vert v_{j,i-1} ,  t_{j,i-1} ; \bm{\theta} ) ) \nonumber \\
& + \frac{1}{2} \frac{\partial^2}{ \partial v^2} ( 2 \theta_t \alpha (v+ p_t) (1 - v- p_t) \, \rho_i(v ,t_{j,i} \vert v_{j,i-1} ,  t_{j,i-1} ; \bm{\theta} ) )\,,  \label{eq:fpk}
\end{align}
subject to the initial conditions $\rho_{i-1}(v , t_{j, i-1} , \bm{\theta} ) = \delta(v - V_{j, t_{i-1}}) \,,$ where $ \delta(v - V_{j, t_{i-1}})$ is the Dirac-delta generalized function centered at $ V_{j, t_{i-1}}\,.$

>>>>

\begin{figure}[h]
  %\includegraphics[width=26mm,scale=1]{}
  \caption{Add Figure showing the difference with and without derivative tracking}
  \label{fig:derivative_tracking}
\end{figure}

@@

Our model (\ref{VtSDE}) for the forecast error has a diffusion term that depends on the state variable $V_t$. To estimate the unknown model parameters, a recommended technique is to modify the SDE (\ref{VtSDE}) by applying the so-called Lamperti transform (see \cite[40--41]{iacus1}, \cite{moma}, \cite[98--100]{saso}) to the process $V$ to obtain a SDE for the transformed process whose diffusion term does not depend anymore on the state variable. To this purpose, we consider the following Lamperti transformation  
%{\color{red} (Remind that continuous-time models with different diffusion term are singular and the likelihood cannot be written down. Besides, it can be unstable to work with low frequency. If the frequency of the data is too low, the Lamperti transform helps to detect this problem (in principle, it is not clear how to determine the right frequency for estimation. )}
%\begin{equation}
%Z_t = h(V_t, t) = \frac{1}{\sqrt{2 \theta_0 \alpha}} \int_{\xi}^{x} \frac{1}{\sqrt{(u + p_t)(1 - u - p_t)}} \,du \Bigg{\vert}_{x = V_t} \,,
%\end{equation}
%where $\xi$ is an arbitrary point of the state space of the process $V$. The choice of $\xi = \frac{1}{2} - p_t$, where $p_t$ is a known input, 
\begin{align}
Z_t = h(V_t, t)  & = \frac{1}{\sqrt{2 \alpha \theta_0 }} \int \frac{1}{\sqrt{(v + p_t)(1 - v - p_t)}} \,dv \Bigg{\vert}_{v = V_t}  \nonumber \\
& = - \sqrt{ \frac{2}{ \alpha \theta_0 }} \arcsin (\sqrt{ 1 - V_t - p_t})
\end{align}
that, after applying It\^{o}'s formula on $h(V_t, t)$, leads to the following SDE with state independent diffusion term
\begin{align}
dZ_t = \Bigg[  & \frac{\dot{p}_t}{ \sqrt{2 \alpha \theta_0 (V_t + p_t)(1 - V_t - p_t)}}  \nonumber \\ 
& + \frac{- \theta_t V_t}{ \sqrt{2  \alpha \theta_0 (V_t + p_t)(1 - V_t - p_t)}} - \frac{1}{4} \frac{\sqrt{2 \alpha \theta_0} \left( 1 - 2 (V_t + p_t)\right)}{\sqrt{(V_t + p_t)(1 - V_t - p_t)}}  \Bigg] \,dt + dW_t \,. \label{eq:stindepSDE}
\end{align}
After replacing $V_t = 1 - p_t - \sin^2 \left(- \sqrt{ \frac{ \alpha \theta_0}{2} } Z_t \right) $ in (\ref{eq:stindepSDE}), we obtain that the  process $Z$ satisfies the SDE
\begin{align}
dZ_t & = \Bigg[ \frac{ \dot{p}_t  - \theta_t  \left(1 - p_t - \sin^2 \Big(- \sqrt{ \frac{ \alpha \theta_0}{2} } Z_t \Big) \right) }{\sqrt{2 \alpha \theta_0} \cos\Big(- \sqrt{ \frac{ \alpha \theta_0}{2} } Z_t \Big)  \sin\Big(- \sqrt{ \frac{ \alpha \theta_0}{2} } Z_t \Big)}  \nonumber \\
&  - \frac{1}{4}  \frac{\sqrt{2 \alpha \theta_0} \left(1 - 2 \cos^2 \Big(- \sqrt{ \frac{ \alpha \theta_0}{2} } Z_t \Big) \right) }{\cos\Big(- \sqrt{ \frac{ \alpha \theta_0}{2} } Z_t \Big)  \sin\Big(- \sqrt{ \frac{ \alpha \theta_0}{2} } Z_t \Big)} \Bigg] \,dt + dW_t \nonumber \\
&  = \left[  \frac{  2  \dot{p}_t - \theta_t (1 - 2 p_t)  + (\alpha \theta_0 - \theta_t) \cos(- \sqrt{2 \alpha \theta_0 } Z_t) }{\sqrt{2 \alpha \theta_0} \sin{(- \sqrt{2 \alpha \theta_0} Z_t)}}  \right] \,dt + dW_t \,. 
\end{align}

@@

Suppose that any of $M$ non-overlapping paths of the continuous-time It\^{o} process $V = \{ V_t, t  \in [0,T] \}$ is sampled at $N + 1$ equispaced discrete points with given length interval $\Delta$, and let $ V^{M,N + 1}=\{ V_{t_1^{N + 1}} , V_{t_2^{N + 1}} ,\ldots , V_{t_M^{N + 1}} \}$ denote this random sample, with $V_{t_j^{N + 1}} = \{ V_{t_j + i \Delta}\,, i = 0, \ldots, N \}, \, j = 1, \ldots, M$.

Let $\rho(v \vert v_{j, i-1} ; \bm{\theta})$ be the conditional probability density of $V_{t_j + i \Delta} \equiv V_{j, i}$ given $V_{j, i-1} = v_{j, i-1}$ evaluated at $v$, where $\bm{\theta} = (\theta_0, \alpha, \epsilon)$ are the unknown model parameters.

The It\^{o} process $V$ defined by the SDE (\ref{VtSDE}) is Markovian, then the likelihood function of the data can be written as the following product of transition densities:  
%Consider the set of $M$ paths with $N$ observations each, $ V^{M,N}=\{ V_{t_1^{M,N}} , V_{t_2^{M,N}} ,\ldots , V_{t_N^{M,N}} \}$, observed in $\Delta_N$ intervals.
\begin{equation}
\mathcal{L}(\bm{\theta}; V^{M,N +1}) = \prod\limits_{j=1}^M \left\{ \prod\limits_{i=1}^N \rho ( {V_{j, i}| V_{j, i-1}} ; p_{[t_{j,  i-1}, t_{j , i} ]},  \bm{\theta} )  \,  \right\} \,,
\label{likelihood}
\end{equation}
where $t_{j ,i} \equiv  t_j + i \Delta$ for any $j = 1, \ldots, M$ and $i = 0, \ldots, N$.

 {\color{red} Remark: The statistical model (\ref{likelihood}) can be generalized incorporating the transition that occurs during the time interval, say $\delta$, between the time when the forecast is done and the first date of forecasting. In this case, the inner product in  (\ref{likelihood}) must include for any of the $M$ paths an additional factor, say $\rho_0 (V_{j, 0}|V_{j, -\delta};\bm{\theta},\delta)$, expressing the conditional density of this early transition. The parameter $\delta$ can be calibrated after the estimation of $\bm{\theta}$, suggesting an optimal time for the scheduling of the forecasts.}

@@

\begin{align}
\frac{ \partial f }{\partial t } & \rho(v ,t \vert v_{j,i-1} ,  t_{j,i-1} ; \bm{\theta} )= - \frac{\partial}{ \partial v} (- \theta_t v \, \rho(v ,t \vert v_{j,i-1} ,  t_{j,i-1} ; \bm{\theta} ) ) \nonumber \\
& + \frac{1}{2} \frac{\partial^2}{ \partial v^2} ( 2 \theta_0 \alpha (v+ p_t) (1 - v- p_t) \, \rho(v ,t \vert v_{j,i-1} ,  t_{j,i-1} ; \bm{\theta} ) )\,,  \label{eq:fpk}
\end{align}
subject to the initial conditions $\rho(v , t_{j, i-1} ; \bm{\theta} ) = \delta(v - V_{j, i-1}) \,,$ where $ \delta(v - V_{j, i-1})$ is the Dirac-delta generalized function centered at $ V_{j, i-1}\,.$
